---
title: "NOAA Analysis"
author: "Zeid El Ameli"
date: "5/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## R Markdown



```{r data, echo=TRUE}
noaa <- read.csv("~/NOAA.csv")
```

## Subseting Our Data

the original dataframe is too large, we don't need all the columns in it. so, we limited the data to only a few columns that we're interested in.

```{r pressure, echo=TRUE}
sub_noaa <- noaa[c("BEGIN_DATE_TIME","END_DATE_TIME","EPISODE_ID","EVENT_ID","STATE","STATE_FIPS","CZ_NAME","CZ_TYPE","CZ_FIPS","EVENT_TYPE","SOURCE","BEGIN_LAT","BEGIN_LON","END_LAT","END_LON")]
head(sub_noaa)
```

## Ordering Our Data

Next step is ordering the data by the beginning year and month.
```{r ordered, echo=TRUE}
arranged_noaa <-noaa[order(noaa$BEGIN_YEARMONTH),]
```

## Changing The Case of States and Counties

In our data, State & County names are stored in upper case. We will now change them to title case.

```{r case_change, echo=TRUE}
noaa$title_state <-str_to_title(noaa$STATE, locale = "en")
noaa$title_cz <-str_to_title(noaa$CZ_NAME, locale = "en")
```

## Limiting Our Data

Now, we want to only include data from CZ_TYPE of "C". 

```{r filtered, echo=TRUE}
filtered <- filter(noaa, CZ_TYPE == "C") %>%
  select(-CZ_TYPE)
```

##Combining Columns

We want to add a "0" in front of our STATE_FIPS & CZ_FIPS columns, then combine the two into 1!

```{r concat, echo=TRUE}
noaa$concat <- str_c("0",noaa$STATE_FIPS,"0",noaa$CZ_FIPS)
```

## Renaming Columns to Lower Case

We noticed that all our column names are in UPPER CASE, we now want to change them to lower case.

```{r tolower, echo=TRUE}
noaa <- as_tibble(noaa)
rename_all(noaa, tolower)
```

##Building Dataframes From Base R Code

Base R has some useful datasets, ones of them is a dataset about US states and some valuable info about each. Let's build a dataframe from that dataset!

```{r states, echo=TRUE}
state_info <- data.frame(state=state.name, region=state.region, area=state.area)
```

## Merging Dataframes

We now want to merge dataframes from both our own imported dataframe, and the base dataframe in R.

```{r join, echo=TRUE}
freq_table <- data.frame(table(noaa$STATE))
freq_table1 <- rename(freq_table,c("state"="Var1"))
freq_table1$title_state <- str_to_title(freq_table1$state, locale = "en")
freq_table1 = freq_table1[-c(1)]
merged <- merge(x=freq_table1,y=state_info,by.x="title_state",by.y="state")
head(merged)
```

## Plotting Our Data

Now, we would like to see a graph that shows us states by area and number of storms hit it in 1994.

```{r plot, echo=TRUE}
plot <- ggplot(merged, aes(x= area, y=Freq)) +
  geom_point(aes(color = region)) +
  labs(x="Land Area (Sq. Miles)", y="Number of Storms in 1994")
plot
```